APPM report - Draft

### 2022-11-08 paper_version/main.tex
minor change

\section{Full Discretization}
\subsection{Motivation: Reduction to 1D Discretization}
Our work is an extension of the 1D scheme proposed in \cite{degond_2012} towards a scheme in three dimensions. Therefore, it is compatible with the original 1D scheme when one considers its reduction to one dimension. This mental experiment aims to help readers understand how the new 3D scheme originates from the 1D one. 
\subsubsection{Projection to one dimension} \label{sec:projection_to_1d}
Consider a mesh pair $(\mathcal{M}, \mathcal{M}')$ that are constructed by 2D extrusion along $z$-axis and are unbounded along transverse directions on $x$-$y$ plane. On assuming the transverse invariance and rotational symmetry, one can reduce the 3D problem to a 1D one along $z$-axis. All the D.o.F. are now defined at the locations that are the projections of their original locations. For instance, an edge that is orthogonal to $z$-axis projects to a vertex while an edge parallel to $z$-axis projects to an edge. The procedure is sketched in Fig. \ref{fig:1d_reduction}.

@@

\subsubsection{1D plasma fully-discrete model} \label{sec:1d_fully_discrete_model}
The following fully-discrete scheme is proposed in \cite{degond_2012}. The spatial discretization can be interpreted by the 1D projection procedure illustrated above. Here we want to discuss the time discretization with a focus on its semi-implicitness. From now on, we attach to variables a superscript involving $m$ to indicate the time step. The scheme is built on a uniform 1D mesh with cell size $h$ and evolves the system with time step $\delta^m := t^{m+1} - t^m$:  
\begin{align*}
    \frac{n^{m+1}_k - n^m_k}{\delta^m} + \frac{\mycolorbox{orange!30}{f^{n,m+1/2}_{k+1/2} - f^{n,m+1/2}_{k-1/2}}}{h} &= 0, \\
    \frac{(nu_x)^{m+1}_k - (nu_x)^m_k}{\delta^m} + \frac{f^{u_x,m}_{k+1/2} - f^{u_x,m}_{k-1/2}}{h} &= -\frac{n^m_k \mycolorbox{orange!30}{E^{m+1}_{x,k}}}{\varepsilon^2} + \frac{n^m_k u^m_{z,k}\Tilde{B}^m_{y,k}}{\varepsilon^2},\\
    \frac{(nu_z)^{m+1}_k - (nu_z)^m_k}{\delta^m} + \frac{f^{u_z,m}_{k+1/2} - f^{u_z,m}_{k-1/2}}{h} &= - \frac{n^m_k \mycolorbox{orange!30}{\Tilde{E}^{m+1}_{z,k}}}{\varepsilon^2} - \frac{n^m_k u^m_{x,k}\Tilde{B}^m_{y,k}}{\varepsilon^2}, \\ 
    \frac{B^{m+1}_{y,k+1/2} - B^{m}_{y,k+1/2}}{\delta^m} + \frac{\mycolorbox{orange!30}{E^{m+1}_{x,k+1} - E^{m+1}_{x,k}}}{h} &= 0, \\
    \lambda^2 \frac{E^{m+1}_{x,k} - E^{m}_{x,k}}{\delta^m} + \frac{\mycolorbox{orange!30}{B^{m+1}_{y,k+1/2} - B^{m+1}_{y,k-1/2}}}{h} &= \mycolorbox{orange!30}{(nu_x)^{m+1}_k}, \\
    \lambda^2 \frac{E^{m+1}_{z,k+1/2} - E^{m}_{z,k+1/2}}{\delta^m} &= \mycolorbox{orange!30}{f^{m+1}_{n,k+1/2}}, 
\end{align*}
where 
\begin{equation*}
    \Tilde{E}^m_{z,k} = \frac{1}{2}(E^m_{z,k+1/2} + E^m_{z,k-1/2}), \ \ \ \Tilde{B}^m_{y,k} = \frac{1}{2}(B^m_{y,k+1/2} + B^m_{y,k-1/2}),
\end{equation*}
are the averaged electromagnetic fields, and the fluxes read
\begin{align*}
    f^{n,m+1/2}_{k+1/2} & = \frac{1}{2}[\mycolorbox{orange!30}{(nu_z)^{m+1}_k + (nu_z)^{m+1}_{k+1}} + \mu^m_{k+1/2}(n^m_k - n^m_{k+1})], \\
    f^{u_x,m}_{k+1/2} & = \frac{1}{2}[(nu_zu_x)^{m}_k + (nu_zu_x)^{m}_{k+1} + \mu^m_{k+1/2}((nu_x)^m_k - (nu_x)^m_{k+1})], \\
    f^{u_z,m}_{k+1/2} & = \frac{1}{2}[(nu_z^2 + p(n)/\varepsilon^2)^{m}_k + (nu_z^2 + p(n)/\varepsilon^2)^{m}_{k+1} + \mu^m_{k+1/2}((nu_z)^m_k - (nu_z)^m_{k+1})].
\end{align*}
The implicit terms are highlighted. The implicitness of the electric field in the Lorentz force is shown to be necessary in \cite{fabre_1992} for the Euler-Poisson system. In \cite{degond_2012}, based on the linearized stability analysis, the author shows that the full-implicit treatment of the Maxwell equations is essential to the stability in the quasi-neutral limit. Besides, the mass flux in the mass conservation equation and the current in Amp\`{e}re's law have the same level of implicitness in order to guarantee the consistency with Gauss's law.

@@

\subsubsection{CFL condition}
It has been shown in \cite{degond_2012} that the stability-induced constraint on the time-step size does not depend on the value of $\lambda$ thanks to the implicitness in the discrete Maxwell system. However, the Euler system is discretized by an explicit scheme and thus still subject to a CFL condition.

@@

\subsubsection{Procedure of time-stepping}
The semi-implicit scheme (\ref{equ:maxwell_faraday_ap})-(\ref{equ:Q_coll_ap}) entails solving a linear system for each time step. We would like to provide a brief guide on how to build such a linear system and evolve the system.

@@

\begin{remark}
    The formula (\ref{equ:discrete_ohm}) functions as Ohm's law in the discrete level and $\mathbb{M}_\sigma^m$ can be treated as a "conductivity" matrix. Since in the quasi-neutral limit $\lambda \rightarrow 0$ we lose control of the $\mathbf{E}$-field through its time derivative, $\mathbb{M}_\sigma^m$ is crucial to the stability of the scheme. Due to the fact that $\mathbf{j}$ and $\mathbf{d}$ are both defined on dual faces, it is reasonable to adopt a local relation between them. Therefore, we perform the \emph{diagonal lumping} trick on (\ref{equ:discrete_ohm}), i.e. rendering it diagonal by accumulating the row sum. We observe enhanced stability in the numerical experiments with very minor changes in the results.
\end{remark}
\section{Numerical Experiment}
We conduct numerical experiments based on the geometric setting and boundary conditions as presented in Section \ref{sec:BC}. The primal-dual mesh pair $(\mathcal{M}, \mathcal{M}')$ is generated based on the code developed in \cite[][ch. 1]{fuchs_2021}. An instance is shown in Fig. \ref{fig:primal_dual_meshes}. We consider a two-fluid plasma model which is composed of electrons and ions. We set the mass ratio $\varepsilon_e := m_e / m_i = 10^4$ and the collision parameter $\alpha^\text{coll} = 32$.
\begin{figure}
    \centering
    \includegraphics[scale=0.3]{primal_mesh.png}
    \hspace{1cm}
    \includegraphics[scale=0.3]{dual_mesh.png}
    \caption{Primal mesh (left) and dual mesh(right).}
    \label{fig:primal_dual_meshes}
\end{figure}
\subsection{Case 1: Quasi-1D Problem}
On noticing the radial symmetry of the domain as well as boundary conditions, we can simplify the governing equations defined at the axis of symmetry, i.e. $z$-axis in this case, thanks to vanishing radial components of the vector fields and their radial derivatives. The Euler-Maxwell system at the axis of symmetry reduces to
\begin{align*}
    &\partial_t n_* + \partial_z(n_*u_{*,z}) = 0, \\
    &\partial_t (n_* u_{*,z}) + \partial_z(n_*u_{*,z}^2) + \varepsilon_*^{-2} \partial_z p_* = \varepsilon_*^{-2}q_*n_*E_z, \\
    &\partial_t (n_*e_*) + \partial_z (n_* h_* u_{*,z}) = \varepsilon_*^{-2} q_* n_* u_{*,z} E_z, \\
    \lambda^2 &\partial_t E_z = \sum_* - q_* n_* u_{*,z}.
\end{align*}
which is similar to the 1D model in Section \ref{sec:projection_to_1d}. Therefore, it is reasonable to compare the results produced by the 1D scheme in Section \ref{sec:1d_fully_discrete_model} and the one on $z$-axis produced by the 3D scheme proposed in Section \ref{sec:3d_ap_scheme}. Fig. \ref{fig:z_axis_reduction} demonstrates good matching. 
\begin{figure}
    \centering
    \includegraphics{z_axis_reduction.eps}
    \caption{Comparison of 1D two-fluid model and reduced 1D model on $z$-axis from 3D two-fluid model. The results are both generated on meshes with 64 cells along $z$-axis. No friction term exists. $t = 5 \times 10^{-4}.$}
    \label{fig:z_axis_reduction}
\end{figure}

\subsection{Case 2: 3D Cylindrical Problem}
We stick to the setting in Section \ref{sec:BC} and use the following initial conditions
\begin{equation*}
    n_\ast(\mathbf{x}, 0) = 1, \ \ \  p_\ast(\mathbf{x},0) = 1, \ \ \forall \mathbf{x}\in\Omega, \ast \in \{\text{electron}, \text{ion}\}.
\end{equation*}
All the other quantities are set zero initially. The electrode at $\Gamma_C^-$(cathode) is grounded, and the one at $\Gamma_C^+$ (anode) has potential 1.0 throughout the simulation, i.e. $\varphi(\mathbf{x}, t)|_{\Gamma_C^-} = 0, \varphi(\mathbf{x}, t)|_{\Gamma_C^+} = 1$. The experiment is intended to investigate how a stationary system is driven by an applied voltage.

@@

\todo{demonstration instability without stabilization}
\todo{redo the mesh study}

>>>>

\section{Full Discretization}
\subsection{Motivation: Reduction to 1D Discretization}
Our work is an extension of the 1D scheme proposed in \cite{degond_2012} towards a scheme in three dimensions. Therefore, it is compatible with the original 1D scheme when one considers its reduction to one dimension. This mental experiment aims to help readers understand how the new 3D scheme originates from the 1D one. 
\subsubsection{Projection to One Dimension} \label{sec:projection_to_1d}
Consider a mesh pair $(\mathcal{M}, \mathcal{M}')$ that are constructed by 2D extrusion along $z$-axis and are unbounded along transverse directions on $x$-$y$ plane. On assuming the transverse invariance and rotational symmetry, one can reduce the 3D problem to a 1D one along $z$-axis. All the D.o.F. are now defined at the locations that are the projections of their original locations. For instance, an edge that is orthogonal to $z$-axis projects to a vertex while an edge parallel to $z$-axis projects to an edge. The procedure is sketched in Fig. \ref{fig:1d_reduction}.

@@

\subsubsection{1D Fully-discrete Plasma Model} \label{sec:1d_fully_discrete_model}
The following fully-discrete scheme is proposed in \cite{degond_2012}. The spatial discretization can be interpreted by the 1D projection procedure illustrated above. Here we want to discuss the time discretization with a focus on its semi-implicitness. From now on, we attach to variables a superscript involving $m$ to indicate the time step. The scheme is built on a uniform 1D mesh with cell size $h$ and evolves the system with time step $\delta^m := t^{m+1} - t^m$:  
\begin{align*}
    \frac{n^{m+1}_k - n^m_k}{\delta^m} + \frac{\mycolorbox{orange!30}{f^{n,m+1/2}_{k+1/2} - f^{n,m+1/2}_{k-1/2}}}{h} &= 0, \\
    \frac{(nu_x)^{m+1}_k - (nu_x)^m_k}{\delta^m} + \frac{f^{u_x,m}_{k+1/2} - f^{u_x,m}_{k-1/2}}{h} &= -\frac{n^m_k \mycolorbox{orange!30}{E^{m+1}_{x,k}}}{\varepsilon^2} + \frac{n^m_k u^m_{z,k}\Tilde{B}^m_{y,k}}{\varepsilon^2},\\
    \frac{(nu_z)^{m+1}_k - (nu_z)^m_k}{\delta^m} + \frac{f^{u_z,m}_{k+1/2} - f^{u_z,m}_{k-1/2}}{h} &= - \frac{n^m_k \mycolorbox{orange!30}{\Tilde{E}^{m+1}_{z,k}}}{\varepsilon^2} - \frac{n^m_k u^m_{x,k}\Tilde{B}^m_{y,k}}{\varepsilon^2}, \\ 
    \frac{B^{m+1}_{y,k+1/2} - B^{m}_{y,k+1/2}}{\delta^m} + \frac{\mycolorbox{orange!30}{E^{m+1}_{x,k+1} - E^{m+1}_{x,k}}}{h} &= 0, \\
    \lambda^2 \frac{E^{m+1}_{x,k} - E^{m}_{x,k}}{\delta^m} + \frac{\mycolorbox{orange!30}{B^{m+1}_{y,k+1/2} - B^{m+1}_{y,k-1/2}}}{h} &= \mycolorbox{orange!30}{(nu_x)^{m+1}_k}, \\
    \lambda^2 \frac{E^{m+1}_{z,k+1/2} - E^{m}_{z,k+1/2}}{\delta^m} &= \mycolorbox{orange!30}{f^{m+1}_{n,k+1/2}}, 
\end{align*}
where 
\begin{equation*}
    \Tilde{E}^m_{z,k} = \frac{1}{2}(E^m_{z,k+1/2} + E^m_{z,k-1/2}), \ \ \ \Tilde{B}^m_{y,k} = \frac{1}{2}(B^m_{y,k+1/2} + B^m_{y,k-1/2}),
\end{equation*}
are the averaged electromagnetic fields, and the fluxes read
\begin{align*}
    f^{n,m+1/2}_{k+1/2} & = \frac{1}{2}[\mycolorbox{orange!30}{(nu_z)^{m+1}_k + (nu_z)^{m+1}_{k+1}} + \mu^m_{k+1/2}(n^m_k - n^m_{k+1})], \\
    f^{u_x,m}_{k+1/2} & = \frac{1}{2}[(nu_zu_x)^{m}_k + (nu_zu_x)^{m}_{k+1} + \mu^m_{k+1/2}((nu_x)^m_k - (nu_x)^m_{k+1})], \\
    f^{u_z,m}_{k+1/2} & = \frac{1}{2}[(nu_z^2 + p(n)/\varepsilon^2)^{m}_k + (nu_z^2 + p(n)/\varepsilon^2)^{m}_{k+1} + \mu^m_{k+1/2}((nu_z)^m_k - (nu_z)^m_{k+1})].
\end{align*}
The implicit terms are highlighted. The implicitness of the electric field in the Lorentz force is shown to be necessary in \cite{fabre_1992} for the Euler-Poisson system. In \cite{degond_2012}, based on the linearized stability analysis, the author shows that the full-implicit treatment of the Maxwell equations is essential to the stability in the quasi-neutral limit. Besides, the mass flux in the mass conservation equation and the current in Amp\`{e}re's law have the same level of implicitness in order to guarantee the consistency with Gauss's law.

@@

\subsubsection{CFL Condition}
It has been shown in \cite{degond_2012} that the stability-induced constraint on the time-step size does not depend on the value of $\lambda$ thanks to the implicitness in the discrete Maxwell system. However, the Euler system is discretized by an explicit scheme and thus still subject to a CFL condition.

@@

\subsubsection{Procedure of Time-stepping}
The semi-implicit scheme (\ref{equ:maxwell_faraday_ap})-(\ref{equ:Q_coll_ap}) entails solving a linear system for each time step. We would like to provide a brief guide on how to build such a linear system and evolve the system.

@@

\begin{remark}
    The formula (\ref{equ:discrete_ohm}) functions as Ohm's law in the discrete level and $\mathbb{M}_\sigma^m$ can be treated as a "conductivity" matrix. Since in the quasi-neutral limit $\lambda \rightarrow 0$ we lose control of the $\mathbf{E}$-field through its time derivative, $\mathbb{M}_\sigma^m$ is crucial to the stability of the scheme. Due to the fact that $\mathbf{j}$ and $\mathbf{d}$ are both defined on dual faces, it is reasonable to adopt a local relation between them. Therefore, we perform the \emph{diagonal lumping} trick on (\ref{equ:discrete_ohm}), i.e. rendering it diagonal by accumulating the row sum. We observe enhanced stability in the numerical experiments with very minor changes in the results.
\end{remark}
\section{Numerical Experiment}

We conduct numerical experiments based on the geometric setting and boundary conditions as presented in Section \ref{sec:BC}. The primal-dual mesh pair $(\mathcal{M}, \mathcal{M}')$ is generated based on the code developed in \cite[][ch. 1]{fuchs_2021}. An instance is shown in Fig. \ref{fig:primal_dual_meshes}. We consider a two-fluid plasma model which is composed of electrons and ions. We set the mass ratio $\varepsilon_e := m_e / m_i = 10^4$ and the collision parameter $\alpha^\text{coll} = 32$.
\begin{figure}
    \centering
    \includegraphics[scale=0.3]{primal_mesh.png}
    \hspace{1cm}
    \includegraphics[scale=0.3]{dual_mesh.png}
    \caption{Primal mesh (left) and dual mesh(right).}
    \label{fig:primal_dual_meshes}
\end{figure}
\subsection{Case 1: Quasi-1D Problem}
On noticing the radial symmetry of the domain as well as boundary conditions, we can simplify the governing equations defined at the axis of symmetry, i.e. $z$-axis in this case, thanks to vanishing radial components of the vector fields and their radial derivatives. The Euler-Maxwell system at the axis of symmetry reduces to
\begin{align*}
    &\partial_t n_* + \partial_z(n_*u_{*,z}) = 0, \\
    &\partial_t (n_* u_{*,z}) + \partial_z(n_*u_{*,z}^2) + \varepsilon_*^{-2} \partial_z p_* = \varepsilon_*^{-2}q_*n_*E_z, \\
    &\partial_t (n_*e_*) + \partial_z (n_* h_* u_{*,z}) = \varepsilon_*^{-2} q_* n_* u_{*,z} E_z, \\
    \lambda^2 &\partial_t E_z = \sum_* - q_* n_* u_{*,z}.
\end{align*}
which is similar to the 1D model in Section \ref{sec:projection_to_1d}. Therefore, it is reasonable to compare the results produced by the 1D scheme in Section \ref{sec:1d_fully_discrete_model} and the one on $z$-axis produced by the 3D scheme proposed in Section \ref{sec:3d_ap_scheme}. Fig. \ref{fig:z_axis_reduction} demonstrates good matching. 
\begin{figure}
    \centering
    \includegraphics{z_axis_reduction.eps}
    \caption{Comparison of 1D two-fluid model and reduced 1D model on $z$-axis from 3D two-fluid model. The results are both generated on meshes with 64 cells along $z$-axis. No friction term exists. $t = 5 \times 10^{-4}.$}
    \label{fig:z_axis_reduction}
\end{figure}

\subsection{Case 2: 3D Cylindrical Problem}

\todo{demonstration instability without stabilization}

\todo{redo the mesh study}

\todo{supplement results for sin voltage}

\todo{visualize in a better way}

We stick to the setting in Section \ref{sec:BC} and use the following initial conditions
\begin{equation*}
    n_\ast(\mathbf{x}, 0) = 1, \ \ \  p_\ast(\mathbf{x},0) = 1, \ \ \forall \mathbf{x}\in\Omega, \ast \in \{\text{electron}, \text{ion}\}.
\end{equation*}
All the other quantities are set zero initially. The electrode at $\Gamma_C^-$(cathode) is grounded, and the one at $\Gamma_C^+$ (anode) has potential 1.0 throughout the simulation, i.e. $\varphi(\mathbf{x}, t)|_{\Gamma_C^-} = 0, \varphi(\mathbf{x}, t)|_{\Gamma_C^+} = 1$. The experiment is intended to investigate how a stationary system is driven by an applied voltage.

@@
